\section{Retrieval-Augmented Generation and Adaptive Agent Memory}

In the context of lifelong learning agents, retrieval-augmented generation (RAG) is conceptualized as an external, non-parametric memory module. The taxonomy of agent memory systems distinguishes between Working Memory (storing active user queries), Episodic Memory (retaining past session logs), Semantic Memory (storing structured domain facts), and Parametric Memory (storing general language patterns inside weights) \parencite{11328884}. Under this classification, RAG serves as an expandable semantic and episodic memory layer, enabling agents to dynamically scale their knowledge boundaries without parameter modification.

\subsection{Selective Forgetting and Psychological Models}
Standard RAG systems treat all documents equally, leading to index bloating and context window clutter. To address this, recent architectures have integrated psychological models of human memory, specifically the Ebbinghaus forgetting curve, to implement selective forgetting and memory consolidation \parencite{Sumida_2025}. 

The LUFY framework \parencite{Sumida_2025} simulates cognitive memory decay by computing importance scores across six psychological dimensions (including retrieval frequency, recency, and emotional arousal). By actively forgetting up to $90\%$ of low-importance memory slots, LUFY increases retrieval precision by $17\%$, prevents context window saturation, and achieves better alignment with human memory judgments.

\subsection{Adapting Conversational Memory to the Legal Domain}
While frameworks like LUFY are designed for conversational chatbots to track user preferences, adapting selective memory to the legal domain introduces unique challenges and opportunities. Unlike conversational logs where importance is highly subjective, legal documents possess objective metadata:
\begin{itemize}
    \item \textbf{Citation Centrality:} The authority of a statute or judicial precedent can be modeled using network analysis (e.g., PageRank or in-degree centrality) on citation graphs.
    \item \textbf{Temporal Validity:} Legal articles have explicit effective dates, amendment timelines, and annulment clauses, requiring strict chronological filtering.
    \item \textbf{Severity and Compliance Risk:} Regulations can be prioritized based on penalty severity or compliance risks.
\end{itemize}
Combining these domain-specific factors into a unified legal importance score is key to optimizing RAG memory in legal LLMs \parencite{nguyen2025vlqacomprehensivelargehighquality}. The main research gap lies in balancing index size reduction (forgetting redundant/outdated regulations) with compliance integrity (ensuring active regulations are never forgotten). This defines a core objective of this study.
