\section{Reason for Choosing the Topic}

In recent years, the intersection of artificial intelligence (AI) and the legal domain, commonly referred to as Legal AI, has emerged as a transformative area of research and practical application. Legal systems worldwide, including that of Vietnam, are characterized by a massive volume of statutory texts, judicial precedents, and administrative documents. Analyzing, interpreting, and applying these rules to specific cases requires rigorous logical deduction, representing a heavy cognitive load for legal professionals and introducing substantial latency in judicial proceedings. Consequently, building automated legal reasoning systems that can assist in legal query answering, statutory rule retrieval, and decision-making has become a priority to enhance the efficiency, consistency, and accessibility of judicial systems \parencite{italiani2025enhancing, le-etal-2025-overview}.

To automate legal reasoning, modern systems heavily rely on state-of-the-art Large Language Models (LLMs). These frontier models have demonstrated remarkable capabilities in complex reasoning, zero-shot statutory interpretation, and generating detailed explanatory rationales. However, deploying commercial, proprietary Large Language Models (such as GPT-4 or Claude) within the legal domain presents severe operational constraints:
\begin{enumerate}
    \item \textbf{Data Privacy and Confidentiality:} Legal queries and case documentations frequently contain highly sensitive personal information, proprietary corporate data, and confidential state records under active investigation. Transmitting such data to external, third-party APIs of commercial providers poses substantial security risks. It often violates national regulations, such as the Cybersecurity Law of Vietnam, and international privacy framework rules (e.g., GDPR), which strictly mandate localized, on-premise data processing and absolute user confidentiality.
    \item \textbf{Prohibitive Operating Costs:} Processing long legal documents and statutory databases via external API calls incurs significant, recurring financial costs. Given the extensive document lengths typical of legal cases, token-based pricing models become financially unsustainable for local courts, law firms, and governmental agencies.
    \item \textbf{Latency and Availability Issues:} API-dependent applications are subject to network bottlenecks, server downtime, and latency spikes, which are unacceptable for real-time applications or high-throughput batch processing pipelines in institutional settings.
\end{enumerate}

To resolve the tension between data security, operational cost, and reasoning performance, deploying lightweight Small Language Models (SLMs) on local, on-premise hardware has emerged as a promising alternative. Models with fewer than three billion parameters (such as the Qwen or Llama families) can be run locally on consumer-grade graphic processing units (GPUs) at minimal cost, ensuring complete data sovereignty and zero external transmission.

Nevertheless, Small Language Models suffer from a severe performance deficit. Due to their restricted parameter capacities, they exhibit weak multi-step logical reasoning, are highly prone to hallucinating incorrect legal articles, and frequently suffer from logic loops. In legal reasoning tasks, a single logical leap or a misinterpretation of a statutory clause can render the entire reasoning chain invalid and lead to an erroneous final verdict. Hence, bridging the "reasoning gap" between large, proprietary models and small, locally deployable models is a critical challenge in Legal AI \parencite{li2026legaldrilldiagnosisdrivensynthesislegal, stacey2024distillingrobustnessnaturallanguage}.

Knowledge distillation (KD) provides a mathematically grounded methodology to address this challenge by transferring the reasoning capabilities of a larger, highly capable model (the teacher) to a lightweight model (the student) \parencite{hinton2015distillingknowledgeneuralnetwork, sanh2020distilbertdistilledversionbert}. In the context of legal reasoning, the teacher model is typically an LLM aligned using advanced reinforcement learning frameworks, such as Group Relative Policy Optimization (GRPO), to generate structured, step-by-step reasoning chains (e.g., Chain-of-Thought). By forcing the student model to mimic the teacher's soft probability distributions (logits) and structured reasoning traces, knowledge distillation allows the student to absorb the "dark knowledge" of the teacher, enhancing its logical consistency while maintaining its compact parameter size.

However, standard knowledge distillation methods encounter limitations in sequential reasoning. Traditional offline logit distillation is bound to a static distribution of gold-standard training sequences (teacher-forcing), which leads to exposure bias during autonomous inference. Furthermore, lightweight models are frequently miscalibrated, preventing them from identifying their own logical blind spots. This research addresses these challenges by proposing a novel three-phase optimization pipeline:
\begin{enumerate}
    \item \textbf{Phase 1: Offline Logit Distillation} establishes a baseline legal vocabulary using a supervised anchor and Quantized Low-Rank Adaptation (QLoRA) \parencite{dettmers2023qloraefficientfinetuningquantized}.
    \item \textbf{Phase 2: On-Policy Knowledge Distillation} mitigates exposure bias by forcing the student to learn from its own generated trajectories, guided by a teacher model operating in an Oracle mode with privileged access to the ground-truth answers \parencite{zhao2026selfdistilledreasoneronpolicyselfdistillation}.
    \item \textbf{Phase 3: Diagnosis-Driven Preference Optimization} performs targeted error correction by utilizing an external Large Language Model as a Judge to diagnose student reasoning errors and align the model via Direct Preference Optimization (DPO) \parencite{rafailov2024directpreferenceoptimizationlanguage}. Crucially, this external evaluation is performed solely at training time using public benchmark datasets; during actual institutional deployment, the aligned student model operates entirely on-premise and locally, ensuring absolute data privacy.
\end{enumerate}
By tailoring this pipeline to the Vietnamese legal domain using representative datasets (including legal natural language inference, multiple-choice questions, and syllogism reasoning), we demonstrate that it is possible to build a lightweight, highly capable, and secure legal assistant. The development of such a model is the core motivation for choosing this research topic.
