\section{Research Subjects and Scope}

To ensure a structured and rigorous evaluation, we define the specific research subjects and operational boundaries of this study.

\subsection{Research Subjects}

The primary subjects analyzed and developed in this research are categorized into model architectures and evaluation datasets:
\begin{enumerate}
    \item \textbf{Small Language Models (SLMs):} We focus on lightweight auto-regressive transformer models with fewer than 3 billion parameters. Specifically, the student architectures evaluated are \texttt{Qwen3-0.6B-Base} and \texttt{Qwen3-1.7B-Base}. These are distilled from a larger, aligned teacher model, \texttt{Qwen3-4B-legal-GRPO}, which serves as the optimization oracle. The teacher model is inherited as a pre-aligned checkpoint from a prior research phase, optimized via Group Relative Policy Optimization on Vietnamese legal text; training this teacher model from scratch is outside the scope of this study.
    \item \textbf{Vietnamese Legal Datasets:} The training and evaluation datasets represent three key legal reasoning tasks in the Vietnamese language:
    \begin{itemize}
        \item \textit{Natural Language Inference (NLI):} Determining the semantic relationship (e.g., \texttt{relevant}, \texttt{not\_relevant}, or \texttt{unknown}) between factual scenarios and target legal hypotheses.
        \item \textit{Multiple-Choice Questions (MCQ):} Answering objective legal queries with options A through D.
        \item \textit{Syllogistic Reasoning:} Generating structured explanations composed of a major premise (statutory rules), a minor premise (case facts application), and a final legal conclusion.
    \end{itemize}
\end{enumerate}

\subsection{Research Scope}

The scope of this study is bounded by the following parameters:
\begin{enumerate}
    \item \textbf{Methodological Boundary:} This research is restricted to post-pretraining adaptation methods. We investigate knowledge distillation (offline logit distillation and on-policy distillation) and direct preference alignment. We do not perform foundational pre-training. Model adjustments are limited to Parameter-Efficient Fine-Tuning (PEFT) techniques, specifically Low-Rank Adaptation (LoRA) and Quantized Low-Rank Adaptation (QLoRA).
    \item \textbf{Domain and Language Boundary:} The investigation is strictly confined to the Vietnamese legal framework and legal terminology. The models are trained and evaluated on Vietnamese statutory documents, administrative codes, and localized judicial queries.
    \item \textbf{Hardware and Operational Constraints:} To reflect the operational realities of local judicial offices and small law firms, all training, distillation, and inference processes are restricted to entry-level datacenter hardware, specifically a single GPU setup with 16GB of Video Random Access Memory (VRAM) using an NVIDIA T4 (Tensor Core GPU), representing a resource-constrained institutional deployment setting.
\end{enumerate}
