\section{Project Plan and Budget}

\subsection{Project Timeline and Implementation Plan}
The proposed research is structured as a 6-month project divided into four sequential phases, progressing from corpus processing to pipeline construction, experimental optimization, and final benchmark verification. The breakdown of tasks, timelines, and milestones is detailed below:

\begin{enumerate}
    \item \textbf{Phase 1: Corpus Acquisition, Normalization, and Indexing (Weeks 1--5)}
    \begin{itemize}
        \item Task 1.1: Crawl and compile Vietnamese statutory codes, decrees, circulars, and Supreme Court judicial precedents.
        \item Task 1.2: Structure documents into discrete paragraphs/clauses and generate initial metadata (validity dates, administrative hierarchy, citation networks).
        \item Task 1.3: Train and evaluate semantic bi-encoder embedding models for dense vector retrieval.
        \item \textit{Milestone 1 (Week 3):} Complete legal corpus compilation and normalization.
        \item \textit{Milestone 2 (Week 5):} Complete dense embedding generation and initial indexing.
    \end{itemize}

    \item \textbf{Phase 2: Core Pipeline Integration (Weeks 6--12)}
    \begin{itemize}
        \item Task 2.1: Establish the RAG retrieval pipeline and configure the compliance hard-gate based on metadata validity rules.
        \item Task 2.2: Implement the Selective Memory module, formulating Ebbinghaus decay functions and legal importance features (citation graph centrality, risk weights).
        \item Task 2.3: Build the P+Q unlearning gate, integration rules for forget requests, and security logging structures.
        \item \textit{Milestone 3 (Week 9):} Functioning baseline RAG with metadata validity filtering.
        \item \textit{Milestone 4 (Week 12):} Deployment of Selective Memory and Unlearning modules.
    \end{itemize}

    \item \textbf{Phase 3: Experimental Optimization and Tuning (Weeks 13--18)}
    \begin{itemize}
        \item Task 3.1: Conduct ablation studies on the selective reranking weights ($w_1, \dots, w_4$) and optimize memory threshold $\theta$.
        \item Task 3.2: Benchmark unlearning effectiveness, universality, and harmlessness on target datasets.
        \item Task 3.3: Run jailbreak simulation tests to measure unlearning robustness against context leakage.
        \item \textit{Milestone 5 (Week 15):} Complete parameter tuning for selective memory reranking.
        \item \textit{Milestone 6 (Week 18):} Finalize unlearning evaluations and robustness check.
    \end{itemize}

    \item \textbf{Phase 4: Benchmark Validation and Thesis Compilation (Weeks 19--24)}
    \begin{itemize}
        \item Task 4.1: Perform comprehensive system validation on Vietnamese legal benchmarks including VLQA \parencite{nguyen2025vlqacomprehensivelargehighquality}, ViLegalNLI \parencite{duong2026vilegalnlinaturallanguageinference}, and LegalSLM \parencite{le-etal-2025-overview}.
        \item Task 4.2: Draft research reports, compile codebase documentation, and perform final code review.
        \item Task 4.3: Write and format the final thesis proposal document.
        \item \textit{Milestone 7 (Week 21):} Achieve benchmark validation target scores.
        \item \textit{Milestone 8 (Week 24):} Submission of the final project report.
    \end{itemize}
\end{enumerate}

\subsection{Risk Assessment and Mitigation Strategies}
Table~\ref{tab:risk_assessment} outlines potential technical and operational risks identified for this project, along with corresponding preventive and mitigation actions.

\begin{table}[H]
\centering
\caption{Project Risk Assessment and Mitigation Matrix.}
\label{tab:risk_assessment}
\begin{tabularx}{\textwidth}{lXX}
\hline
\textbf{Identified Risk} & \textbf{Impact \& Probability} & \textbf{Mitigation Strategy} \\
\hline
Metadata Inaccuracy & High Impact / Moderate Probability & Implement a double-pass manual verification check on a randomized sample of 5\% of the ingested legal documents. \\
Retrieval Recall Failure & High Impact / Low Probability & Integrate hybrid search (combining dense bi-encoder with sparse BM25) to act as a fallback retrieval mechanism. \\
Compute Bottleneck & Moderate Impact / Moderate Probability & Pre-compute dense embeddings offline and restrict hyperparameter tuning search grids via Bayesian optimization. \\
Access Restriction Leakage & High Impact / Low Probability & Conduct automated validation audits on the access-control filter policy before deploying index updates. \\
\hline
\end{tabularx}
\end{table}

\subsection{Estimated Research Budget}
The estimated budget required to complete the 6-month research plan is structured under standard institutional categories. The total projected cost is 1,250,000,000 VND (~$50,000$ USD). A detailed breakdown of costs is presented in Table~\ref{tab:estimated_budget}.

\begin{table}[H]
\centering
\caption{Estimated Budget Breakdown for the 6-Month Research Project.}
\label{tab:estimated_budget}
\begin{tabularx}{\textwidth}{lXXr}
\hline
\textbf{Category} & \textbf{Description} & \textbf{Quantity / Rate} & \textbf{Cost (VND)} \\
\hline
\textbf{Personnel} & Principal Investigator (PI) & 6 months @ 25,000,000 VND/mo & 150,000,000 \\
                   & Lead NLP Researcher & 6 months @ 40,000,000 VND/mo & 240,000,000 \\
                   & Data Annotation Assistants (2) & 4 months @ 15,000,000 VND/mo each & 120,000,000 \\
\hline
\textbf{Compute \& Cloud} & GPU Server Rental (A100 instances) & 200 hours @ 80,000 VND/hour & 16,000,000 \\
                         & Vector Database Host (Pinecone/Milvus) & 6 months @ 5,000,000 VND/mo & 30,000,000 \\
                         & Local Development Station (GPU-equipped) & 1 Unit (RTX 4090 workstation) & 110,000,000 \\
\hline
\textbf{Data \& Operations} & Law Corpus License \& Cleaning Fees & Lump sum & 250,000,000 \\
                            & Annotation Tooling \& QC Platform & Licenses & 40,000,000 \\
                            & Legal Expert Validation Fees & 5 Experts @ 15,000,000 VND each & 75,000,000 \\
\hline
\textbf{Administrative} & Publication Fees & 2 IEEE Journal / Conference papers & 80,000,000 \\
                        & Workspace, Travel, \& Communication & General operational costs & 39,000,000 \\
                        & Institutional Indirect Costs (Overhead) & 8\% of direct research costs & 100,000,000 \\
\hline
\textbf{Total Budget} & & & \textbf{1,250,000,000} \\
\hline
\end{tabularx}
\end{table}
