\chapter*{CONTINUAL LEARNING AND CATASTROPHIC FORGETTING IN LARGE LANGUAGE MODELS\\ABSTRACT}

Large Language Models (LLMs) have demonstrated exceptional reasoning capabilities across diverse domains, including the complex field of legal question-answering and logical inference. However, maintaining the temporal validity of their knowledge remains a critical challenge due to the dynamic nature of legislative updates, regulatory amendments, and confidentiality restrictions. Traditional continual learning methodologies that modify model weights directly through Continual Pre-Training (CPT) or Continual Instruction Tuning (CIT) are severely limited by catastrophic forgetting and parameter weight drift. Furthermore, executing targeted machine unlearning (such as removing repealed statutes or confidential records) from neural network weights is computationally expensive and difficult to verify.

This seminar report presents a comprehensive literature review and proposed framework for managing continual learning and catastrophic forgetting in Vietnamese legal LLMs. We systematically analyze current mitigation strategies, ranging from rehearsal, regularization, and parameter-efficient fine-tuning (PEFT/LoRA) to semi-parametric retrievable gradients (ReGrad). We evaluate the advantages of a RAG-centric non-parametric approach, which decouples reasoning from fact retrieval by keeping the base model parameters frozen. Specifically, we detail a selective RAG memory system utilizing Ebbinghaus-based consolidation and decay to optimize vector storage and reduce context noise, coupled with a compliance hard-gate ensuring that valid legal clauses are never omitted. Additionally, we formalize a controllable RAG-based unlearning gate operating under a dual Retriever Policy ($P$) and Prompt Constraint ($Q$) optimization framework to handle temporal, restricted access, and personal data deletion requests. 

The proposed framework is domain-general and is designed to be validated on Vietnamese legal reasoning benchmarks, including VLQA, ViLegalNLI, and LegalSLM, with Vietnamese labour law and tax law adopted as two representative testbed sectors---the former as the primary testbed and the latter as a focused secondary task through which sequential cross-sector catastrophic forgetting is measured. Our review shows that RAG-centric selective memory and behavioral unlearning provide a highly secure, compliant, and cost-efficient methodology for deploying legal assistant systems in low-resource settings, offering zero-weight-drift guarantees and instant knowledge updates compared to resource-intensive parametric fine-tuning alternatives.
